---
title: "COVID-19 Epidemiological Surveillance"
subtitle: "Synthetic Dashboard — Synthea Massachusetts"
format: 
  dashboard:
    orientation: rows
    theme: cosmo
    css: dashboard.css
    nav-buttons:
      - icon: github
        href: "https://github.com/mmaturano-epidemio/covid_surveillance_dashboard"
embed-resources: true
---
  
```{r}
#| label: setup
#| include: false

pacman::p_load(data.table, echarts4r, here, htmltools)

source(here("scripts", "1_setup.R"))
source(here("scripts", "3_demographics.R"))
source(here("scripts", "4_spacial_distribution.R"))
source(here("scripts", "5_temporal_distribution.R"))
```

# Overview

```{=html}
<div class="hero">
  <div class="hero-content">
    <h1>COVID-19 Epidemiological Surveillance</h1>
    <p>An interactive portfolio dashboard analyzing demographic, spatial, and temporal trends in Massachusetts using the synthetic Synthea™ dataset.</p>
    <a href="https://github.com/mmaturano-epidemio/covid_surveillance_dashboard" target="_blank" class="hero-btn">View Repository</a>
  </div>
</div>
```

  
# Demographics

## Row 

### Column {width="50%"}

#### Row {height="100%"}


```{r}
#| title: "Population pyramid"
pyramid
```

### Column {width="35%"}

#### Row {height="100%"}


```{r}
#| title: "Distribution by race"
race
```


### Column {width="35%"}

#### Row {height="50%"}
  
```{r}
#| content: valuebox
#| title: "Male % — overall"
list(
  icon = "gender-male",
  color = "primary",
  value = paste0(male_pct_general, "%")
)
```

```{r}
#| content: valuebox
#| title: "Male % — hospitalized"
list(
  icon = "gender-male",
  color = "info",
  value = paste0(male_pct_hospitalized, "%")
)
```

```{r}
#| content: valuebox
#| title: "Male % — deceased"
list(
  icon = "gender-male",
  color = "danger",
  value = paste0(male_pct_deceased, "%")
)
```

#### Row {height="50%"}

```{r}
#| content: valuebox
#| title: "Mean age — overall"
list(
  icon = "calendar3",
  color = "primary",
  value = mean_age_general
)
```

```{r}
#| content: valuebox
#| title: "Mean age — hospitalized"
list(
  icon = "calendar3",
  color = "info",
  value = mean_age_hospitalized
)
```

```{r}
#| content: valuebox
#| title: "Mean age — deceased"
list(
  icon = "calendar3",
  color = "danger",
  value = mean_age_deceased
)
```

# Spatial Distribution

## Row

### Column {width="40%"}

```{r}
#| title: "COVID-19 incidence"
county_table
```

### Column {width="60%"}

```{r}
#| title: "Cases and diagnosis centers"
map
```


# County performance

## Row

### Column {width="30%"}
```{r}
#| title: "Hospitalizations by county"
top_counties
```

### Column {width="35%"}


```{r}
#| title: "Hospital days by county"

county_average_days

```

### Column {width="35%"}
```{r}
#| title: "Lethality by county"

county_lethality
```


# Temporal Trends

```{r}
#| title: "Patients by county"
temporal_chart
```

# About

## Row

### Column {width="60%"}

#### **Project Overview**
This dashboard demonstrates an end-to-end epidemiological surveillance workflow — 
from relational database extraction to interactive geospatial and temporal 
analysis — built entirely on **synthetic patient data**. It uses the [Synthea™](https://synthetichealth.github.io/synthea/) COVID-19 dataset to model realistic public health metrics in Massachusetts.

#### *Data Source*
All records were generated using [Synthea](https://github.com/synthetichealth/synthea), 
an open-source synthetic patient population simulator developed by MITRE. 
No real patient, institution, or geographic incidence pattern is represented — 
hospital names and case counts are fictional.

### Column {width="40%"}

#### **Technical Stack & Methodology**
The project relies on a highly reproducible and efficient pipeline:

* **Data Manipulation:** Heavy lifting and data wrangling performed entirely with R's `data.table` for high-performance memory management and execution speed.
* **Visualization:** Interactive, modular charts built with `echarts4r` and custom CSS styling.
* **Framework:** Compiled as a standalone Quarto Dashboard, utilizing the `here` package for robust, relative path management.

